---
title: "Exploratory graphics"
output: html_document
date: "2026-01-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(data.table)
library(ggplot2)
theme_set(theme_minimal()) # personal preference - feel free to change!
```

```{r}
# CHANGE THIS to match the data path on your computer
datadir <- "~/data/msk_chord_2024/data/"

# Load the data
clin_patient <- fread(paste0(datadir, "data_clinical_patient.txt"), skip = 4)
clin_sample  <- fread(paste0(datadir, "data_clinical_sample.txt"), skip = 4)
muta <- fread(paste0(datadir, "data_mutations.txt"))
cna  <- fread(paste0(datadir, "data_cna.txt"))
sv   <- fread(paste0(datadir, "data_sv.txt"))
```

# What do we have?

Sample-level information
```{r}
colnames(clin_sample)
```

Patient-level information
```{r}
colnames(clin_patient)
```

# Example: mutation count per cancer type
```{r}

muta_with_patient <- merge(
    muta,
    clin_sample[, .(SAMPLE_ID, PATIENT_ID, CANCER_TYPE)],
    by.x = "Tumor_Sample_Barcode",
    by.y = "SAMPLE_ID",
    all.x = TRUE
  )

forplot <- muta_with_patient[, .N, by = .(Hugo_Symbol, CANCER_TYPE)]
ggplot(forplot, aes(x = CANCER_TYPE, y = N)) +
  geom_boxplot(outliers = FALSE) +
  coord_flip()
```

# Example: Mutations heatmap 
Show top 10 mutated genes × samples, colored by mutation type. Facet or group by cancer type.
(Maybe just for a subset of patients!)

# Example: Co-mutation plot
For a given cancer type, show which gene pairs tend to be mutated together vs mutually exclusive. (Maybe a matrix of odds ratios or correlation coefficients?)

# Example: Copy number aberration comparison
Pick two samples with different CNA patterns and plot them stacked for comparison. 
Maybe: Add gene annotations for known oncogenes/tumor suppressors at their chromosomal positions.

# Example: Treatment timeline visualization
For a single patient, create a timeline showing: diagnosis, treatments (as horizontal bars), progression events (as points), and when sequencing occurred.

# Example: Metastatic site patterns
Which tumor sites co-occur? Create a heatmap or network showing how often pairs of metastatic sites appear together in the same patient.

# Example: Survival curves by genomic features
Create Kaplan-Meier curves comparing patients with/without a specific mutation, stratified by cancer type. 

# Mutation signatures over chromosomes
For one cancer type, plot mutation density across the genome (binned). Are there mutational "hotspots"?

